---
title: "A3: Incarceration"
output: html_document
---
```{r, echo=FALSE} 
source("summary.R")
```
### Introduction
In this report I will be examining the data set of the prison population rate per 100,000 people starting in 1990. I will compare the average prison population rates of different races and genders. This is important because people of color are incarcerated at disproportionately high rates compared to white people, which is considered to be the result of both historical and present-day systemic racism. It is important to examine the trends and comparisons seen in this data to possibly help identify areas that need more attention.

### Summary Information

While examining the data, I could see that it shows that people of color are incarcerated at significantly higher rates than white people, with black people being incarcerated the most. 
The average population of white prisoners per 100,000 is 
`r mean_white_prison_pop_rate_function` . 
The average population of Latinx prisoners per 100,000 is 
`r mean_latinx_prison_pop_rate_function`. 
The average population of AAPI prisoners per 100,000 is 
`r mean_aapi_prison_pop_rate_function`. 
The average population of native prisoners per 100,000 is 
`r mean_native_prison_pop_rate_function`.  
The average population of black prisoners per 100,000 is 
`r mean_black_prison_pop_rate_function`.

```{r, echo=FALSE} 
source("summary.R")

```

### The Dataset

The data was collected by the Vera Institute of Justice

How was the data collected or generated?  
Why was the data collected?  
How many observations (rows) are in your data?  
How many features (columns) are in the data?  
What, if any, ethical questions or questions of power do you need to consider when working with this data?  
What are possible limitations or problems with this data?   (at least 200 words)


### Trends Over Time Chart

This is a chart comparing the mean prison populations of white men and black men. I included this chart in order to highlight racial inequalities of incarceration rates. As you can see, there was a large spike in the prison population around 1997, but black male prison rates are significantly higher than white male prison rates.


```{r, echo = FALSE, code = readLines("TrendOverTime.R")}
```

### Variable Comparison Chart

This is a line graph that compares the prison population of black men and black women. I included this chart to show that black men have been imprisoned significantly more than black women over time. There seems to be a large spike in the prison population around the year 1995 in both demographics, with a much larger spike in the black male prison population.

```{r, echo = FALSE, code = readLines("ComparisonGraph.R")}
```

### Map

Include a chart. Make sure to describe why you included the chart, and what patterns emerged

The last chart that you will create and include will show how a variable is distributed geographically. Again, think carefully about what such a comparison means and what you want to communicate to your user (you may have to find relevant trends in the dataset first!). Here are some requirements to help guide your design:

- Your map needs a title
- Your color scale needs a legend with a clear label
- Use a map based coordinate system to set the aspect ratio of your map

